---
title: "Lab 14 HW"
output: 
  html_document: 
    keep_md: yes
author: "your name here"
date: '2022-02-28'
---


### Load the Libraries
```{r}
library(tidyverse)
library(janitor)
library("palmerpenguins")
library(skimr)

#install.packages("ggVennDiagram")
library(ggVennDiagram)
library(RColorBrewer)

#install.packages("ggworldcloud")
library(ggwordcloud)

options(warn = -1)

options(scipen=999) #cancels the use of scientific notation for the session
```


### Data
```{r}
superhero_info <- readr::read_csv("data/heroes_information.csv", na = c("", "-99", "-"))
superhero_powers <- readr::read_csv("data/super_hero_powers.csv", na = c("", "-99", "-"))

beachbugs_long <- readr::read_csv("data/beachbugs_long.csv")
```

### 1. 
Clean up the column names (no capitals, not spaces) of `superhero_info`, then use 2 functions to remind yourself of structure of the `superhero_info` data set.

```{r}
superhero_info <- superhero_info %>%
  clean_names()
skim(superhero_info)
head(superhero_info, n=10)
```

### 2.
Are bad guys bigger? Make box-plots of weight by `alignment` (alignment on the x-axis).
```{r}
superhero_info %>%
  ggplot(mapping = aes(x=alignment, y=weight, fill=alignment))+
  geom_boxplot()+
  labs(title = "Superhero Weight", 
       x="Alignment", 
       y="Weight")+
  theme(plot.title=element_text(hjust=.5))
```

### 3. 
Now, make a violin plot of weight by `alignment` (alignment on the x-axis). Add some color!
  What information can you observe in the violin plot that was not visible in the boxplot?
```{r}
superhero_info %>%
  filter(!is.na(weight))%>%
  ggplot(mapping = aes(x=alignment, y=weight, fill=alignment))+
  geom_violin()+
  coord_flip()+
  labs(title = "Superhero Weight", 
       x="Alignment", 
       y="Weight")+
  theme(plot.title=element_text(hjust=.5))
```

#### The violin plots all extend toaward the big value of weight, so extreme values occur for all alignment. The group _alignment == good_ does not give a standard violin plot. The group _alignment == NA_ has a really abnormal plot and I do not know what happened. 

### 4. 
Use `alpha = .5` in `geom_boxplot()` and `geom_violin()` to layer both plots on top of one another. What does this tell you about the distribution of weight in "`bad`" guys?
```{r}
superhero_info %>%
  filter(!is.na(weight))%>%
  ggplot(mapping = aes(x=alignment, y=weight, fill=alignment))+
  geom_boxplot(alpha = .2)+
  geom_violin(alpha = .6)+
  coord_flip()+
  labs(title = "Superhero Weight", 
       x="Alignment", 
       y="Weight")+
  theme(plot.title=element_text(hjust=.5))
```

#### Both boxplot and violin indicate that the weight of guys of bad alignment is mainly around 80. This distributioin is highly concentrated around 80. 

### 5. 
Box plots are great for showing how the distribution of a numeric variable (e.g. weight) varies among a categorical variable (e.g. alignment).
  Make your own violin plot with the superhero data. 
  What is your numeric data? 
  What is your categorical variable?

```{r}
superhero_info%>%
  filter(!is.na(height))%>%
  ggplot(mapping = aes(x=gender, y=height, fill=gender))+
  geom_boxplot(alpha = .2)+
  geom_violin(alpha = .6)+
  coord_flip()+
  labs(title = "Superhero Height", 
       x="Gender", 
       y="Height")+
  theme(plot.title=element_text(hjust=.5))
```

#### The numeric data is height and the categorical data is sex. 

### 6. 
Remind yourself what `beachbugs` looks like. Then generate a heatmap of buglevels by site and year. 
color it with `scale_fill_gradient(low="yellow", high="red")` or colors of your choice. you may find it looks best with one color!
(dont forget, `coord_flip()` is a quick way to improve the look of your plot if you dont like the default orientation)

```{r}
beachbugs_long <- beachbugs_long %>%
  clean_names()
skim(beachbugs_long)
```

```{r}
beachbugs_long%>%
  ggplot(aes(site, year, fill=buglevels))+
  geom_tile()+
  scale_fill_gradient(low="yellow", high="red")+
  theme(axis.text.x = element_text(angle = 60, hjust = 1))+
  labs(title = "Bug level", 
       x="Site", 
       y="Year")+
  theme(plot.title=element_text(hjust=.5))
```

### 7.  
Use the provided code to normalize the beachbug data set. 
Then make a heatmap with the `beachbugs_normalized` data, and use the same color chois as above. Which heatmap do you think is more informative? why?

```{r}
#makes a new column of the highest buglevel observed at each site
beachbugs_w_max <- beachbugs_long %>% 
  group_by(site) %>% 
  mutate(max_buglevel = max(buglevels, na.rm=T)) %>% 
  arrange(site)
beachbugs_w_max

#makes a new table where the buglevels are normalized to the max_buglevel
beachbugs_normalized <- beachbugs_w_max %>% 
  group_by(site) %>% 
  mutate(norm_buglevel = buglevels/max_buglevel) %>% 
  arrange(site,year) %>%
  select(site, year, norm_buglevel)   # you dont have to select(), but I think its a clearer looking table
beachbugs_normalized
```
 
```{r}
beachbugs_normalized%>%
  ggplot(aes(site, year, fill=norm_buglevel))+
  geom_tile()+
  scale_fill_gradient(low="yellow", high="red")+
  theme(axis.text.x = element_text(angle = 60, hjust = 1))+
  labs(title = "Bug level", 
       x="Site", 
       y="Year")+
  theme(plot.title=element_text(hjust=.5))
```

#### The later one is more informative because the later one makes a horizontal comparison among different regions to indicate the relative bug level instead of a absolute one. 

### 8.
Let's make a venn diagram of `superhero_info`, from 4 questions:
Is their alignment evil? 
Are their eyes red?
Are they male?
Are they bald?

Start by making the 4 vectors, then the list of vectors. The vector for alignment is provided:
### super heros venn
```{r}
# evil

evil_vec <- superhero_info %>%
  filter(alignment == "bad")%>%
  pull(name)

# red eyes

red_vec <- superhero_info %>%
  filter(eye_color == "red")%>%
  pull(name)

# male

male_vec <- superhero_info %>%
  filter(gender == "Male")%>%
  pull(name)

# bald

bald_vec <- superhero_info %>%
  filter(hair_color == "No Hair")%>%
  pull(name)

```

Your list of vectors will look something like this:
```{r}
questions_list <- list(evil_vec, red_vec, male_vec, bald_vec)

```

### 9. 
Let's make the venn diagram! use the code from lab as a reference. 
```{r}
ggVennDiagram(questions_list , category.names = c("Evil", "Red eye", "Male", "Bald"))
```


### 10. Choose one intersection of the venn diagram that is interesting to you. Use dplyr to find the names of the superheros in that intersection. 

```{r}
superhero_info%>%
  filter(hair_color == "No Hair")%>%
  filter(gender == "Male")%>%
  filter(eye_color == "red")%>%
  filter(alignment != "bad")
```

#### EMMMMM, Why there is no Joel Ledford????? Is he an evil guy :p

### 11. Make another venn diagram with the `superhero_info` data. What are your questions? ( At least 2!) 
```{r}
# Evil

evil_vec <- superhero_info %>%
  filter(alignment == "bad")%>%
  pull(name)

# female

female_vec <- superhero_info %>%
  filter(gender == "Female")%>%
  pull(name)

# blond

blond_vec <- superhero_info %>%
  filter(hair_color == "Blond")%>%
  pull(name)

interest_list <- list(evil_vec, female_vec, blond_vec)
ggVennDiagram(interest_list , category.names = c("Evil", "Female", "Blond"))
```

#### This is for attractive evil lady. 

### 12.
What are some very common super powers? Lets make a word cloud with the `superhero_powers` data.

Use the provided code to make the frequency table, then make a word cloud with it. 
Remember, you can change `scale_size_area(max_size = 20)` to a different number if the words wont fit!

```{r}
head(superhero_powers)
```
```{r}
power_frequency <- superhero_powers %>% 
  select(-hero_names) %>%
  summarise_all(~(sum(.))) %>%
  pivot_longer(everything(), names_to = "power", values_to = "freq")
power_frequency
```


```{r}
power_frequency%>%
  arrange(desc(freq))%>%
  head(n=120)%>%
  ggplot(aes(
  label = power, 
  size = freq,
  color = power
  )) +
  geom_text_wordcloud() +
  scale_size_area(max_size=10) +
  theme_minimal()
```

### 13.  
Who are some very powerful supers? 
 Lets make a different word cloud with the `superhero_powers` data. 
Use the provided code to make the frequency table, then make a word cloud with it.
You can use `hero_names` for the labels, and `sum_powers` for size. 

```{r}

power_quantity <- superhero_powers %>% 
  pivot_longer(-hero_names, names_to = "power", values_to = "yes_or_no") %>% 
  group_by(hero_names) %>% 
  mutate(sum_powers = sum(yes_or_no, na.rm=T)) %>%
  arrange(desc(sum_powers), hero_names, desc(yes_or_no))
power_quantity

#we have to trim down to only the top 50, or it will crowd the image!
power_quantity <- power_quantity %>% 
  ungroup %>%
  distinct(hero_names, sum_powers) %>%
  slice_max(sum_powers, n = 50)
power_quantity

```



```{r}
power_quantity%>%
  ggplot(aes(
  label = hero_names, 
  size = sum_powers,
  color = hero_names
  )) +
  geom_text_wordcloud() +
  scale_size_area() +
  theme_minimal()
```

## That's it! 🎉
Thanks for coding with us all winter! 
Now go finish your group project :) 

