---
title: "Lab 3 Homework"
author: "Hongtao Qian (Steve)"
date: "`r Sys.Date()`"
output:
  html_document: 
    keep_md: yes
    theme: spacelab
---

## Instructions
Answer the following questions and complete the exercises in RMarkdown. Please embed all of your code and push your final work to your repository. Your final lab report should be organized, clean, and run free from errors. Remember, you must remove the `#` for the included code chunks to run. Be sure to add your name to the author header above.  

Make sure to use the formatting conventions of RMarkdown to make your report neat and clean!  

## Load the tidyverse
```{r message=FALSE, warning=FALSE}
library(tidyverse)
```

## Mammals Sleep
1. For this assignment, we are going to use built-in data on mammal sleep patterns. From which publication are these data taken from? Since the data are built-in you can use the help function in R.
```{r}
??sleep
?msleep
```

#### The source of this dataset is from V. M. Savage and G. B. West. A quantitative, theoretical framework for understanding mammalian sleep. Proceedings of the National Academy of Sciences, 104 (3):1051-1056, 2007.

2. Store these data into a new data frame `sleep`.
```{r}
sleep <-msleep
sleep
```

3. What are the dimensions of this data frame (variables and observations)? How do you know? Please show the *code* that you used to determine this below.  
```{r}
dim(sleep)
str(sleep)
```

4. Are there any NAs in the data? How did you determine this? Please show your code.  
```{r}
is.na.data.frame(sleep)
```

#### Based on this code, I can see a bunch of logical value organized in a new data frame. All 'true' stands for Na in the original data. 

5. Show a list of the column names is this data frame.
```{r}
names(sleep)
```

6. How many herbivores are represented in the data?  
```{r}
herbi<-subset(sleep, vore=="herbi")
  nrow(herbi)
```

7. We are interested in two groups; small and large mammals. Let's define small as less than or equal to 1kg body weight and large as greater than or equal to 200kg body weight. Make two new dataframes (large and small) based on these parameters.
```{r}
large<-subset(sleep, bodywt>=200)
large

small<-subset(sleep, bodywt<=1)
small
```

8. What is the mean weight for both the small and large mammals?
```{r}
small_mean<-small$bodywt
mean(small_mean)
```

```{r}
large_mean<-large$bodywt
mean(large_mean)
```

9. Using a similar approach as above, do large or small animals sleep longer on average?  
```{r}
small_average<-small$sleep_total
mean(small_average)
```

```{r}
large_average<-large$sleep_total
mean(large_average)
```
#### Smaller animals should sleep longer. 

10. Which animal is the sleepiest among the entire dataframe?
```{r}
least<-sleep
least<-arrange(least, sleep_total)
head(least, n=5)
```
#### Giraffe seems to sleep the least. 

```{r}
most<-sleep
most<-arrange(most, desc(sleep_total))
head(most, n=5)
```
#### Little brown bat seems to sleep the most. 


## Push your final code to GitHub!
Please be sure that you check the `keep md` file in the knit preferences.   